# Data Cleaning

```{r}
horses<-read.csv("../../../data/00-raw-data/horses.csv")
performances<-read.csv("../../../data/00-raw-data/performances.csv")
```

## Let's start by cleaning horses:
```{r}
head(horses,2)
```


We are going to remove "horse_name_zh" which contains the horses' names in Cantonese,
"horse_url", "owner_url", "sire", "no_of_start_past_10_meetings", "dam","dam_sire", "same_sire and "last_updated" which are all irrelevant to our analysis
```{r}
horses.clean<-horses[,-c(3,4,13,16,19,20,22,23,24,25,26)]
head(horses.clean,2)
```

We are only interested in horses that run in Hong Kong while there are some horses
in this data set that raced in Singapore.
```{r}
library(dplyr)
horses.clean<-horses.clean%>%
  filter(horse_country == "HK")
```

Now we want to clean "no_of_1_2_3". This column represents how many 1st, 2nd, and 3rd place
finishes the horse has ever had. The last number is the total number of races the horse
has participated in. We are going to break up this column into seperate columns
for placing finishes and number of races.
```{r}
library(stringr)

split1 <-str_split_fixed(horses.clean$no_of_1_2_3, "-", 4)
horses.clean$first_place <- split1[,1]
horses.clean$second_place <- split1[,2]
horses.clean$third_place <- split1[,3]
horses.clean$races <- split1[,4]
# Remove original column
horses.clean <- horses.clean[,-15]
head(horses.clean,2)
```


## Now let's clean Performances
```{r}
head(performances,2)
```
We are going to remove the column containing the Cantonese name for the horse,
jockey name in Cantonese, jockey url, trainer name in Cantonese, trainer url,
race url, running positions (redundant information),and last updated, and length behind winner. Length behind winner being deleted is okay because we have LBW for all 4 sections in another variable.
```{r}
perf.clean<-performances[,-c(3,15,16,19,20,23,28,33,37)]
head(perf.clean,2)
```


We are going to once again filter out any race in the dataset that happened in 
Singapore.
```{r}
perf.clean<-perf.clean%>%
  filter(race_country == "HK")
```

Let's get rid of race_country column now that all of the values should equal "HK"
```{r}
perf.clean<-perf.clean%>%
  select(-race_country)
```

Now I am going to do one of the important parts of data cleaning for this analyis:
Breaking the section data into 4 distinct columns, both for LBW and time.

```{r}
# Getting rid of brackets and commas in the columns
perf.clean$sections <- str_replace_all(perf.clean$sections,"\\[|\\]|\\,|\\{","")
perf.clean$sections <- str_replace_all(perf.clean$sections,"\\}",",")
perf.clean$sections_time <- str_replace_all(perf.clean$sections_time,"\\[|\\]|\\,|\\{","")
perf.clean$sections_time <- str_replace_all(perf.clean$sections_time,"\\}",",")
```


```{r}
# splitting each section into their own column
split2 <-str_split_fixed(perf.clean$sections,",",6)
perf.clean$section1<-split2[,1]
perf.clean$section2<-split2[,2]
perf.clean$section3<-split2[,3]
perf.clean$section4<-split2[,4]
perf.clean$section5<-split2[,5]
perf.clean$section6<-split2[,6]
```

```{r}
# same thing for section time
split3 <-str_split_fixed(perf.clean$sections_time,",",6)
perf.clean$section1_time<-split3[,1]
perf.clean$section2_time<-split3[,2]
perf.clean$section3_time<-split3[,3]
perf.clean$section4_time<-split3[,4]
perf.clean$section5_time<-split3[,5]
perf.clean$section6_time<-split3[,6]
```

Getting rid of our original section and section_time columns for redundancy reasons.
```{r}
perf.clean<-select(perf.clean,-c(sections,sections_time))
```

Removing characters from section time column and converting to numeric
```{r}
library(mgsub)
perf.clean$section1_time <- mgsub(perf.clean$section1_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section2_time <- mgsub(perf.clean$section2_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section3_time <- mgsub(perf.clean$section3_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section4_time <- mgsub(perf.clean$section4_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section5_time <- mgsub(perf.clean$section5_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section6_time <- mgsub(perf.clean$section6_time,c('time','"',':',','),
                                  c('','','',''))
```

Saving the section time columns as numeric, along with our finishing time column.
```{r}
perf.clean$section1_time <- as.numeric(perf.clean$section1_time)
perf.clean$section2_time <- as.numeric(perf.clean$section2_time)
perf.clean$section3_time <- as.numeric(perf.clean$section3_time)
perf.clean$section4_time <- as.numeric(perf.clean$section4_time)
perf.clean$section5_time <- as.numeric(perf.clean$section5_time)
perf.clean$section6_time <- as.numeric(perf.clean$section6_time)
```
```{r}
perf.clean$finish_time <- rowSums(perf.clean[,c("section1_time","section2_time",
                                                "section3_time","section4_time",
                                                "section5_time","section6_time")],na.rm=TRUE)
```


```{r}
head(perf.clean,2)
```

Now I want to create a variable that is a measure of a horse's past performance. A past performance metric could be created to be a very complex variable that weights a multitude of features such as increases in rating over time, historical placement, finishing time history, etc.. I am going to simply create a variable that is the average placement of the horse up until the race date. I could do average race finishing time, but due to every horse running at different distances, I would have to create a an average finishing time for a horse for each distance up until the race date. This method could get computationally ineffcient and also really clutter our data with high dimensionality. For this reason, I am just going to do average placement up until the race date.

```{r}
# Convert race_date to a Date object
perf.clean$race_date <- as.Date(perf.clean$race_date)

# Arrange the data by horse_id and race_date
perf.clean <- perf.clean %>%
  arrange(horse_id, race_date)

perf.clean <- perf.clean %>%
  group_by(horse_id) %>%
  mutate(average_placement = lag(cummean(final_placing))) %>%
  ungroup()

# For the first race of each horse, set average_placement to NA
perf.clean$average_placement[!duplicated(perf.clean$horse_id)] <- NA

# View the updated data frame
head(perf.clean)
```
Now I also want to create a variable that indicates how much rest a horse was able to get between races.

```{r}
perf.clean <- perf.clean %>%
  group_by(horse_id) %>%
  mutate(days_between = c(0, diff(race_date)))

head(perf.clean)
```

Export cleaned data to csv
```{r}
write.csv(horses.clean,"../../../data/01-modified-data/horses.clean.csv",row.names = FALSE)
```

```{r}
write.csv(perf.clean,"../../../data/01-modified-data/perf.clean.csv",row.names = FALSE)
```


