# Data Cleaning

```{r}
horses<-read.csv("../../../data/00-raw-data/horses.csv")
performances<-read.csv("../../../data/00-raw-data/performances.csv")
```

## Let's start by cleaning horses:
```{r}
head(horses,2)
```


We are going to remove "horse_name_zh" which contains the horses' names in Cantonese,
"horse_url", "owner_url", "sire", "no_of_start_past_10_meetings", "dam","dam_sire", "same_sire and "last_updated". *Note for grader: I will go into more depth why I am removing these in the future*
completion of this project.
```{r}
horses.clean<-horses[,-c(3,4,13,16,19,20,22,23,24,25,26)]
head(horses.clean,2)
```

We are only interested in horses that run in Hong Kong while there are some horses
in this data set that raced in Singapore.
```{r}
library(dplyr)
horses.clean<-horses.clean%>%
  filter(horse_country == "HK")
```

Now we want to clean "no_of_1_2_3". This column represents how many 1st, 2nd, and 3rd place
finishes the horse has ever had. The last number is the total number of races the horse
has participated in. We are going to break up this column into seperate columns
for placing finishes and number of races.
```{r}
library(stringr)

split1 <-str_split_fixed(horses.clean$no_of_1_2_3, "-", 4)
horses.clean$first_place <- split1[,1]
horses.clean$second_place <- split1[,2]
horses.clean$third_place <- split1[,3]
horses.clean$races <- split1[,4]
# Remove original column
horses.clean <- horses.clean[,-15]
head(horses.clean,2)
```


## Now let's clean Performances
```{r}
head(performances,2)
```
We are going to remove the column containing the Cantonese name for the horse,
the date (as the data became corrupted through the original API and it is not
necessarily important when we have the year of the season and measurements of 
the track quality to account for weather seasonality impacts on the track),
jockey name in Cantonese, jockey url, trainer name in Cantonese, trainer url,
race url, running positions (redundant information),and last updated, and length behind winner. Length behind winner being deleted is okay because we have LBW for all 4 sections in another variable.
```{r}
perf.clean<-performances[,-c(3,7,15,16,19,20,23,28,33,37)]
head(perf.clean,2)
```


We are going to once again filter out any race in the dataset that happened in 
Singapore.
```{r}
perf.clean<-perf.clean%>%
  filter(race_country == "HK")
```

Let's get rid of race_country column now that all of the values should equal "HK"
```{r}
perf.clean<-perf.clean%>%
  select(-race_country)
```

Now I am going to do one of the important parts of data cleaning for this analyis:
Breaking the section data into 4 distinct columns, both for LBW and time.

```{r}
# Getting rid of brackets and commas in the columns
perf.clean$sections <- str_replace_all(perf.clean$sections,"\\[|\\]|\\,|\\{","")
perf.clean$sections <- str_replace_all(perf.clean$sections,"\\}",",")
perf.clean$sections_time <- str_replace_all(perf.clean$sections_time,"\\[|\\]|\\,|\\{","")
perf.clean$sections_time <- str_replace_all(perf.clean$sections_time,"\\}",",")
```


```{r}
# splitting each section into their own column
split2 <-str_split_fixed(perf.clean$sections,",",6)
perf.clean$section1<-split2[,1]
perf.clean$section2<-split2[,2]
perf.clean$section3<-split2[,3]
perf.clean$section4<-split2[,4]
perf.clean$section5<-split2[,5]
perf.clean$section6<-split2[,6]
```

```{r}
# same thing for section time
split3 <-str_split_fixed(perf.clean$sections_time,",",6)
perf.clean$section1_time<-split3[,1]
perf.clean$section2_time<-split3[,2]
perf.clean$section3_time<-split3[,3]
perf.clean$section4_time<-split3[,4]
perf.clean$section5_time<-split3[,5]
perf.clean$section6_time<-split3[,6]
```

Getting rid of our original section and section_time columns for redundancy reasons.
```{r}
perf.clean<-select(perf.clean,-c(sections,sections_time))
```

Removing characters from section time column and converting to numeric
```{r}
library(mgsub)
perf.clean$section1_time <- mgsub(perf.clean$section1_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section2_time <- mgsub(perf.clean$section2_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section3_time <- mgsub(perf.clean$section3_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section4_time <- mgsub(perf.clean$section4_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section5_time <- mgsub(perf.clean$section5_time,c('time','"',':',','),
                                  c('','','',''))
perf.clean$section6_time <- mgsub(perf.clean$section6_time,c('time','"',':',','),
                                  c('','','',''))
```

Saving the section time columns as numeric, along with our finishing time column.
```{r}
perf.clean$section1_time <- as.numeric(perf.clean$section1_time)
perf.clean$section2_time <- as.numeric(perf.clean$section2_time)
perf.clean$section3_time <- as.numeric(perf.clean$section3_time)
perf.clean$section4_time <- as.numeric(perf.clean$section4_time)
perf.clean$section5_time <- as.numeric(perf.clean$section5_time)
perf.clean$section6_time <- as.numeric(perf.clean$section6_time)
```
```{r}
perf.clean$finish_time <- rowSums(perf.clean[,c("section1_time","section2_time",
                                                "section3_time","section4_time",
                                                "section5_time","section6_time")],na.rm=TRUE)
```


```{r}
head(perf.clean,2)
```
Export cleaned data to csv
```{r}
write.csv(horses.clean,"../../../data/01-modified-data/horses.clean.csv",row.names = FALSE)
```

```{r}
write.csv(perf.clean,"../../../data/01-modified-data/perf.clean.csv",row.names = FALSE)
```


